---
title: "Indeed Product Scientist Homework"
author: "Michael Lee"
date: "September 1, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Indeed sales leads

The provided dataset on sales leads contained over 70,000 observations and 9 variables. There were many missing data fields that needed to be addressed prior to analysis. For example, empty fields in the "revenue" column appear as "NA". First, the dataset was read into R-studio and the variables were reviewed. I then replaced the "NA" fields with "0" values.

```{r}
dataset <- read.csv("C:/Users/cwbishop/Documents/Github/Indeed/homework_data_set.csv")
names(dataset)
dataset$revenue[is.na(dataset$revenue)]=0
```

## Question 1
"How many leads are represented in this dataset? Please describe both the assigned and unassigned populations.
What is the average revenue of each group?"

Each lead is uniquely numbered and the dataset is ordered according to the lead number. If the data numbering were perfect, one could use the last number (77890) as a count of the number of leads. However, since the first lead starts on "0" and to account for potential data entry errors, a count of the number of rows would be more appropriate.

```{r}
nleads <- nrow(dataset)
print(nleads)
```
To count the number of leads that were assigned or unassigned, we can use the table function on the dataset grouped by the "assigned" variable. Here, I am assuming a "1" in the assigned column is an assigned lead.

```{r}
nleads_assign <- table(dataset$assigned)
names(nleads_assign) <- c("Not assigned","Assigned")
print(nleads_assign)
```

To further describe the assigned and unassigned population, we first subset the dataset by the "assigned" variable

```{r}
data_assigned <- dataset[dataset$assigned==1,]
data_unassigned <- dataset[dataset$assigned==0,]
```

To calculate the average revenue in each group, I use the aggregate-mean function on the dataset, aggregating by the assigned variable. Since I am averaging revenue, it is critical to assign "0" to the missing revenue fields, which I assume represents no revenue was generated for that lead. This is important because one group may have more overall revenue due to more leads, which does not provide insight into the impact of assignment. 

```{r}
avg_revenue <- as.numeric(aggregate(revenue~assigned,dataset,FUN=mean)[2][,1])
names(avg_revenue) <- c("Unassigned","Assigned")

barplot(avg_revenue,main="Average Revenue by assignment", xlab="Assignment", ylab="$ dollars")
```

The above plot shows that average revenue is higher in the assigned group than unassigned.

## Question 2
"What are the most important metrics to consider in this data set? Why?"

There are 9 variables in the dataset. The most important variables in regard to the questions asked are 1) "revenue", the dependent variable (outcome), 2) "assigned" will tell us if assigning leads is effective, and 3) "age" will be neccessary to determine a relationship between account age and revenue.

## Question 3
"Analyze any existing relationship between account age and revenue, through visualization and/or other means."

Plotting Revenue by age, it is difficult to visually assess a relationship because a majority of the leads result in 0 revenue, and there are over 70,000 data points in the dataset. The first plot shows revenue by age of account. 

```{r}
plot(dataset$age,dataset$revenue, main="Revenue by age of account", ylab="Revenue", xlab="Age (days")
dataset <- dataset[order(dataset$age),]
dataset$rev_cumsum <- cumsum(dataset$revenue)
```

To better assess the relationship between account age and revenue, I converted revenue into a cumulative sum ordered by age. Here, we can see that almost a third of the revenue-producing leads occur very close to the date they are created. The leads over time continue to lead to revenue until about 2000 days, where the revenue plateaus. 

```{r}
plot(dataset$age,dataset$rev_cumsum, main="Cumulative revenue by age of account",ylab="Cumulative Revenue", xlab="Age (days)")
```

##Question 4
"What is the incremental value from assigning a lead to the sales team?"

To examine the incremental value from assigning a lead to the sales team, I used a linear model with a single independent variable.

```{r}
rev_mod <- lm(dataset$revenue~dataset$assigned)
summary(rev_mod)
```

Assignment to a sales team has a significant effect on revenue. The coefficient estimate shows that assignment results in a 2,199,845 increase.  

To account for other potential confounders, a linear model was used with revenue as the outcome, and the other variables in the dataset as independent variables. 

```{r}
all_mod <- lm(dataset$revenue~dataset$advertiser_id+dataset$assigned+as.integer(dataset$date_assignment_starts)+as.integer(dataset$date_assignment_ends)+as.integer(dataset$first_revenue_date)+as.integer(dataset$date_created)+dataset$age+dataset$assign_days)
summary(all_mod)
```

After accounting for the other variables, the estimate for assignment is 1,158,000 compared to no assignment and is statistically significant. Other variables are significant as well, and warrant further investigation.

##Bonus Question
"Investigate the data however you like and discuss any interesting insights you can find in the data"

A major potential confounding factor is described as "potential value". The highest probability of a sale is assigned to the sales rep. Without knowing this value, it is difficult to know the impact of the sale reps. An assigned lead had a higher probability of a sale before assignment. To investigate further, I subsetted the dataset by assignment and plotted the cumulative sum revenue by the first revenue date. All assignments occurred between 2/1/2017 and 2/3/2017.

```{r}

dataset <- dataset[order(dataset$first_revenue_date),]
dataset$rev_cumsum <- cumsum(dataset$revenue)

data_unassigned <- data_unassigned[order(data_unassigned$first_revenue_date),]
data_unassigned$rev_cumsum <- cumsum(data_unassigned$revenue)

data_assigned <- data_assigned[order(data_assigned$first_revenue_date),]
data_assigned$rev_cumsum <- cumsum(data_assigned$revenue)

plot(dataset$first_revenue_date,dataset$rev_cumsum, main="Cumulative revenue by first revenue date",xlab="First Revenue Date", ylab="Cumulative Revenue")
lines(data_unassigned$first_revenue_date,data_unassigned$rev_cumsum, col='red')
lines(data_assigned$first_revenue_date,data_assigned$rev_cumsum, col='blue')
legend(0,100000000000,c("Total cumulative revenue","Assigned cumulative revenue","Unassigned cumulative revenue"), lty=c(1,1), lwd=c(2.5,2.5,2.5), col=c("black","blue","red"))

```


This plot shows several interesting points: 1) There is a large increase in sales near 2/1/2017, both in assigned and unassigned leads, 2) the leads that were eventually assigned had sales prior to being assigned, while unassigned leads did not and 3) there was more revenue in unassigned leads than assigned on 2/1/2017. It is interesting that there were sales of unassigned leads on 2/1/2017. This is unexpected, because there should be no reason unassigned leads would produce revenue on that date. The assigned leads that led to sales before assignment may be due to the higher probability of sale by an unaccounted factor. The smaller increase in assigned leads than unassigned may be due to the smaller pool of available leads in the assigned group because some of the leads have already converted to a sale. 

